---
title: "Campus Greenspace and Chronic Absenteeism"
subtitle: "California K-12 Public Schools, 2022"
author: "Logan Peters · [GitHub](https://github.com/loganpeters97) · [Email](mailto:loganapeters1@gmail.com)"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
---
<style>
.main-container { padding-bottom: 0 !important; margin-bottom: 0 !important; }
body { margin-bottom: 0 !important; padding-bottom: 0 !important; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(readr); library(dplyr); library(tidyr); library(stringr)
library(ggplot2); library(gt); library(sandwich); library(lmtest)
library(sf); library(tigris); library(tidycensus)
options(tigris_use_cache = TRUE)
```

# 1. Keys and Functions

```{r keys}
## CDS key handling.
pad14 <- function(x) str_pad(str_remove(as.character(x), "\\.0+$"), 14, pad = "0")
pad11 <- function(x) str_pad(as.character(x), 11, pad = "0")

## collapse messy multi-line CDE headers (newlines -> spaces, squish)
clean_names <- function(df){ names(df) <- str_squish(str_replace_all(names(df), "\n", " ")); df }

## strips commas, %, stray chars
parse_val <- function(x) as.numeric(str_replace_all(as.character(x), "[^0-9.\\-]", ""))

## builds a 14-digit CDSCode from county/district/school code columns (CDE files)
make_cds <- function(df) df %>%
  mutate(CDSCode = pad14(paste0(
    str_pad(`County Code`, 2, pad="0"),
    str_pad(`District Code`, 5, pad="0"),
    str_pad(`School Code`, 7, pad="0"))))

check_cds <- function(df, name) {
  n <- dplyr::n_distinct(df$CDSCode)
}
```

# 2. Read Data  
### California Department of Education Outcomes, Greenspace Metrics, Census Socioeconmic Measures and Composite SES Index Creation

```{r read}
## ---- outcome + covariates (CDE files, keyed by county/district/school) ----
absent <- clean_names(read_csv("/Users/logan/Practicum/Campuses/chronicabsenteeism22-v3.csv"))
frpm   <- clean_names(read_csv("/Users/logan/Practicum/Campuses/frpm2122_v2.csv"))

##  BUFFER EXPOSURES (centroid buffers, 500m & 1000m) ----
ndvi_evi_buf <- read_csv("/Users/logan/Practicum/Greenspace Constructs/csds2025_schools_stacked_centroids_ndvi_evi_annual.csv",
                         col_types = cols(CDS_uniq = col_character()))
canopy_buf   <- read_csv("/Users/logan/Practicum/Greenspace Constructs/csds2025_schools_stacked_centroids_urban_tree_canopy_cover.csv",
                         col_types = cols(CDS_uniq = col_character()))

## CAMPUS-LEVEL EXPOSURES (school polygons, no buffer) ----
ndvi_evi_campus <- read_csv("/Users/logan/Practicum/Greenspace Constructs/csds2025_schools_stacked_ndvi_evi_annual.csv",
                            col_types = cols(CDS_uniq = col_character()))
canopy_campus   <- read_csv("/Users/logan/Practicum/Greenspace Constructs/csds2025_schools_stacked_urban_tree_canopy_cover.csv",
                            col_types = cols(CDS_uniq = col_character()))


## ---- Building the tract-level SES index from ACS (same as PLACES) 
## Pull the component variables, compute proportions, z-standardize, average (unemployment reverse-signed) so higher SES_INDEX = higher socioeconomic status.
acs_raw <- get_acs(
  geography = "tract", state = "CA", year = 2023, survey = "acs5",
  variables = c(
    MHI          = "B19013_001",   # median household income
    MHV          = "B25077_001",   # median home value (owner-occupied)
    EDU_TOTAL    = "B15003_001",   # pop 25+ (education denominator)
    BA           = "B15003_022",
    MA           = "B15003_023",
    PROF         = "B15003_024",
    PHD          = "B15003_025",
    INTDIV_TOTAL = "B19054_001",   # households (interest/div denominator)
    INTDIV_YES   = "B19054_002",
    HOUSING_TOTAL= "B25002_001",   # total housing units
    OCCUPIED     = "B25002_002",
    LABOR_FORCE  = "B23025_003",   #labor force 16+
    UNEMPLOYED   = "B23025_005"),
  output = "wide", geometry = FALSE) %>%
  rename_with(~ str_remove(., "E$"), ends_with("E")) %>%   # drop the estimate "E" suffix
  mutate(GEOID = pad11(GEOID))

## proportions
ses_components <- acs_raw %>%
  transmute(
    GEOID,
    MHI, MHV,
    EDUCATION   = ifelse(EDU_TOTAL     > 0, (BA + MA + PROF + PHD) / EDU_TOTAL, NA_real_),
    PCT_INTDIV  = ifelse(INTDIV_TOTAL  > 0, INTDIV_YES / INTDIV_TOTAL,          NA_real_),
    PCT_OCCUPIED= ifelse(HOUSING_TOTAL > 0, OCCUPIED / HOUSING_TOTAL,           NA_real_),
    PCT_UNEMPLOYED = ifelse(LABOR_FORCE > 0, UNEMPLOYED / LABOR_FORCE,          NA_real_))

## z-standardize each component; reverse-sign unemployment so higher = higher SES
z <- function(x) (x - mean(x, na.rm = TRUE)) / sd(x, na.rm = TRUE)

ses_index <- ses_components %>%
  mutate(
    z_MHI          = z(MHI),
    z_MHV          = z(MHV),
    z_EDUCATION    = z(EDUCATION),
    z_PCT_INTDIV   = z(PCT_INTDIV),
    z_PCT_OCCUPIED = z(PCT_OCCUPIED),
    z_PCT_UNEMPLOYED = -z(PCT_UNEMPLOYED)) %>%          # reverse-signed
  rowwise() %>%
  mutate(SES_INDEX = mean(c(z_MHI, z_MHV, z_EDUCATION, z_PCT_INTDIV,
                            z_PCT_OCCUPIED, z_PCT_UNEMPLOYED), na.rm = TRUE)) %>%
  ungroup() %>%
  select(GEOID, SES_INDEX) %>%
  filter(!is.na(SES_INDEX)) %>%
  distinct(GEOID, .keep_all = TRUE)

```

# 3. Buffer Exposure Dataframe (500m & 1000m in one df)

```{r buffer-exposures}

gs_year <- 2022   # analysis year for NDVI/EVI

## NDVI/EVI: rows are CDS x buffer x year. Keep the analysis year, pivot the two buffers (feature_buffer_m = 500, 1000) into separate columns per index
ndvi_evi_wide <- ndvi_evi_buf %>%
  mutate(CDSCode = pad14(CDS_uniq)) %>%
  filter(year == gs_year, feature_buffer_m %in% c(500, 1000)) %>%
  group_by(CDSCode, feature_buffer_m) %>%
  summarise(ndvi_annual_mean = mean(ndvi_annual_mean, na.rm = TRUE),
            evi_annual_mean  = mean(evi_annual_mean,  na.rm = TRUE),
            .groups = "drop") %>%
  pivot_wider(names_from = feature_buffer_m,
              values_from = c(ndvi_annual_mean, evi_annual_mean),
              names_glue = "{.value}_{feature_buffer_m}") %>%
  rename(ndvi_500  = ndvi_annual_mean_500,  ndvi_1000 = ndvi_annual_mean_1000,
         evi_500   = evi_annual_mean_500,   evi_1000  = evi_annual_mean_1000)

## Canopy: rows are CDS x radius (500, 1000) x year-label. The file contains
## canopy_2022 AND canopy_2018; keep 2022 to match the NDVI/EVI analysis year.
## Pivot canopy_pct by radius.
canopy_wide <- canopy_buf %>%
  mutate(CDSCode = pad14(CDS_uniq)) %>%
  filter(radius_m %in% c(500, 1000), label == "canopy_2022") %>%
  group_by(CDSCode, radius_m) %>%
  summarise(canopy_pct = mean(canopy_pct, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = radius_m, values_from = canopy_pct,
              names_glue = "canopy_{radius_m}")

## one buffer-exposure row per school
buffer_exposures <- full_join(ndvi_evi_wide, canopy_wide, by = "CDSCode")

## ---- campus-level exposures ----
## NDVI/EVI at the school polygon, 2022.
ndvi_evi_campus_wide <- ndvi_evi_campus %>%
  mutate(CDSCode = pad14(CDS_uniq)) %>%
  filter(year == gs_year) %>%
  group_by(CDSCode) %>%
  summarise(ndvi_campus = mean(ndvi_annual_mean, na.rm = TRUE),
            evi_campus  = mean(evi_annual_mean,  na.rm = TRUE),
            .groups = "drop")

## canopy at the school polygons (keep 2022 label)
canopy_campus_wide <- canopy_campus %>%
  mutate(CDSCode = pad14(CDS_uniq)) %>%
  filter(label == "canopy_2022") %>%
  group_by(CDSCode) %>%
  summarise(canopy_campus = mean(canopy_pct, na.rm = TRUE), .groups = "drop")

## merge campus-level onto the buffer frame -> polygons + 500m + 1000m per measure
buffer_exposures <- buffer_exposures %>%
  full_join(ndvi_evi_campus_wide, by = "CDSCode") %>%
  full_join(canopy_campus_wide,   by = "CDSCode")

## guard: confirm the CDS codes are intact (not truncated)
check_cds(buffer_exposures, "buffer_exposures")
```

# 4. Analytic Dataframe

```{r covariates}
## chronic absenteeism: school totals only (Aggregate Level S, Reporting Cat TA)
absent_sel <- absent %>%
  make_cds() %>%
  filter(`Aggregate Level` == "S", `Reporting Category` == "TA") %>%
  transmute(CDSCode,
            chronic_rate   = parse_val(ChronicAbsenteeismRate),
            chronic_enroll = parse_val(ChronicAbsenteeismEligibleCumulativeEnrollment))

frpm_sel <- frpm %>% make_cds() %>%
  transmute(CDSCode,
            enroll_k12   = parse_val(`Enrollment (K-12)`),
            pct_frpm_k12 = parse_val(`Percent (%) Eligible FRPM (K-12)`),
            charter      = dplyr::recode(str_squish(`Charter School (Y/N)`),
                                         "Y" = "Yes", "N" = "No", .default = NA_character_))
```

```{r school-tract-link}
## Link each school to its census tract (for the tract-level SES index and for SE clustering). Uses the CSCD campus geodatabase.
gdb_path <- "/Users/logan/Practicum/Campuses/CSCD_2025.gdb"
campus <- st_read(gdb_path, layer = "Schools_Current_Stacked") %>%
  mutate(CDSCode = pad14(CDSCode)) %>% st_make_valid() %>% filter(!st_is_empty(.))

tracts_ca <- tigris::tracts(state = "CA", year = 2022, cb = TRUE) %>%
  select(GEOID) %>% st_transform(st_crs(campus))

school_tract <- campus %>% select(CDSCode) %>% st_point_on_surface() %>%
  st_join(tracts_ca, join = st_intersects, left = TRUE) %>%
  st_drop_geometry() %>% distinct(CDSCode, .keep_all = TRUE) %>%
  mutate(GEOID = pad11(GEOID))

school_grade <- campus %>%
  st_drop_geometry() %>%
  mutate(CDSCode = pad14(CDSCode),
         level_raw = as.character(Level)) %>%
  distinct(CDSCode, .keep_all = TRUE) %>%
  mutate(grade_level = dplyr::recode(level_raw,
    "Elementary"                                               = "Elementary",
    "Intermediate/Middle/Junior High"                          = "Middle",
    "High School"                                              = "High",
    "Elementary - High Combination"                            = "Combination",
    "Elementary - Intermediate/Middle/Junior High Combination" = "Combination",
    "Intermediate/Middle/Junior High - High Combination"       = "Combination",
    .default = NA_character_),
    grade_level = factor(grade_level,
                         levels = c("Elementary","Middle","High","Combination"))) %>%
  select(CDSCode, grade_level)
```

```{r assemble}
analysis <- school_tract %>%
  left_join(frpm_sel,        by = "CDSCode") %>%
  left_join(absent_sel,      by = "CDSCode") %>%
  left_join(buffer_exposures,by = "CDSCode") %>%
  left_join(school_grade,    by = "CDSCode") %>%   # grade band for stratification
  left_join(ses_index,       by = "GEOID")     # tract SES index attached to school

## IQR-scale each buffer exposure so coefficients are per-IQR and comparable
iqr1 <- function(x) x / IQR(x, na.rm = TRUE)
analysis <- analysis %>% mutate(
  ndvi_campus_iqr = iqr1(ndvi_campus), evi_campus_iqr = iqr1(evi_campus),
  canopy_campus_iqr = iqr1(canopy_campus),
  ndvi_500_iqr   = iqr1(ndvi_500),   ndvi_1000_iqr   = iqr1(ndvi_1000),
  evi_500_iqr    = iqr1(evi_500),    evi_1000_iqr    = iqr1(evi_1000),
  canopy_500_iqr = iqr1(canopy_500), canopy_1000_iqr = iqr1(canopy_1000),
  log_enroll = ifelse(enroll_k12 > 0, log(enroll_k12), NA_real_),
  charter    = factor(charter, levels = c("No","Yes")))

## SES quartiles (higher = higher SES), defined over tracts then carried to schools
ses_q_tbl <- analysis %>% distinct(GEOID, SES_INDEX) %>%
  filter(!is.na(SES_INDEX)) %>%
  mutate(ses_q = dplyr::ntile(SES_INDEX, 4),
         ses_q = factor(ses_q, labels = c("Q1 (lowest SES)","Q2","Q3","Q4 (highest SES)"))) %>%
  select(GEOID, ses_q)
analysis <- analysis %>% left_join(ses_q_tbl, by = "GEOID")
```

# 5. Main models (per exposure x buffer)

```{r ols-helper}
covars <- c("pct_frpm_k12", "log_enroll", "charter")

## OLS with tract-clustered SEs; returns tidy coefficient row for the exposure
fit_ols <- function(exposure, label, data = ols_df) {
  need <- c("chronic_rate", exposure, covars, "GEOID")
  d <- data %>% filter(if_all(all_of(need), ~ !is.na(.)))
  m <- lm(reformulate(c(exposure, covars), "chronic_rate"), data = d)
  V <- vcovCL(m, cluster = d$GEOID)
  b <- coef(m)[exposure]; se <- sqrt(diag(V))[exposure]
  data.frame(Exposure = label, beta = b,
             ci_low = b - 1.96*se, ci_high = b + 1.96*se,
             p = 2*pnorm(-abs(b/se)), N = nrow(d), row.names = NULL)
}

ols_df <- analysis %>%
  filter(charter %in% c("Yes","No"), !is.na(chronic_rate))
## Exclude small schools where absenteeism rates are unstable.
min_elig <- 100

ols_df <- ols_df %>% filter(chronic_enroll >= min_elig)

nrow(ols_df)   # confirm new analytic N

## exposure -> label map: campus polygons + both buffers, all three measures
exp_map <- c(
  ndvi_campus_iqr   = "NDVI Campus",  ndvi_500_iqr    = "NDVI 500m",    ndvi_1000_iqr   = "NDVI 1000m",
  evi_campus_iqr    = "EVI Campus",   evi_500_iqr     = "EVI 500m",     evi_1000_iqr    = "EVI 1000m",
  canopy_campus_iqr = "Canopy Campus",canopy_500_iqr  = "Canopy 500m",  canopy_1000_iqr = "Canopy 1000m")

ols_results <- do.call(rbind, lapply(names(exp_map), function(e)
  fit_ols(e, exp_map[[e]])))
```

### Descriptives
```{r}
## campus sample desc
nrow(ols_df)                                          # analytic N
ols_df %>% count(grade_level)                         # grade distribution
ols_df %>% count(charter)                             # charter Yes/No
summary(ols_df$chronic_rate)                          # absenteeism center/spread
sd(ols_df$chronic_rate, na.rm = TRUE)
sapply(ols_df[c("ndvi_campus","canopy_campus")], function(x)
  c(min=min(x,na.rm=T), max=max(x,na.rm=T), IQR=IQR(x,na.rm=T)))  # exposure ranges


#attempt 1
## summary statistics (post-exclusion sample, removed schools with less than 100 students) 
n_tot <- nrow(ols_df)
pct   <- function(cond) sprintf("%s (%.1f)", format(sum(cond, na.rm=TRUE), big.mark=","), 100*mean(cond, na.rm=TRUE))
medq  <- function(x, d=1) sprintf(paste0("%.",d,"f (%.",d,"f\u2013%.",d,"f)"),
                                  median(x,na.rm=TRUE), quantile(x,.25,na.rm=TRUE), quantile(x,.75,na.rm=TRUE))

## two characteristic|value blocks side by side (5 rows instead of 10)
campus_t1 <- tibble::tibble(
  C1 = c("Elementary", "Middle", "High", "Combination", "Charter"),
  V1 = c(pct(ols_df$grade_level=="Elementary"), pct(ols_df$grade_level=="Middle"),
         pct(ols_df$grade_level=="High"), pct(ols_df$grade_level=="Combination"),
         pct(ols_df$charter=="Yes")),
  C2 = c("Absenteeism, mean (SD)", "Absenteeism, median (IQR)",
         "NDVI, median (IQR)", "Tree canopy %, median (IQR)", "FRPM %, median (IQR)"),
  V2 = c(sprintf("%.1f (%.1f)", mean(ols_df$chronic_rate,na.rm=TRUE), sd(ols_df$chronic_rate,na.rm=TRUE)),
         medq(ols_df$chronic_rate), medq(ols_df$ndvi_campus, 2),
         medq(ols_df$canopy_campus, 1), medq(ols_df$pct_frpm_k12, 1))
)

campus_t1 %>%
  gt() %>%
  tab_header(title = md(paste0("Campus sample characteristics (*N* = ",
                               format(n_tot, big.mark=","), " California K\u201312 schools, 2022)"))) %>%
  cols_label(C1="", V1="", C2="", V2="") %>%
  tab_style(style = cell_text(weight = "bold"), locations = cells_body(columns = c(C1, C2))) %>%
  cols_align("left", columns = c(C1, C2)) %>%
  cols_align("right", columns = c(V1, V2)) %>%
  tab_source_note(md("*n (%) unless noted. Excludes 683 schools with <100 absenteeism-eligible students.*")) %>%
  tab_options(
    table.font.size = px(10),
    heading.title.font.size = px(11),
    heading.border.bottom.width = px(1.5),
    heading.border.bottom.color = "black",
    column_labels.border.top.style = "none",
    column_labels.border.bottom.style = "none",
    column_labels.padding = px(0),
    table_body.hlines.style = "none",
    table.border.top.style = "none",
    table.border.bottom.width = px(1.5),
    table.border.bottom.color = "black",
    data_row.padding = px(2),
    source_notes.font.size = px(8))

#adjusted to match tract summary table, breaking into two tables
n_tot <- nrow(ols_df)

cat_t1 <- tibble::tibble(
  Characteristic = c("Elementary", "Middle", "High", "Combination",
                     "Charter", "Traditional public"),
  Group = c(rep("Grade level", 4), rep("School type", 2)),
  n = c(sum(ols_df$grade_level=="Elementary", na.rm=TRUE),
        sum(ols_df$grade_level=="Middle", na.rm=TRUE),
        sum(ols_df$grade_level=="High", na.rm=TRUE),
        sum(ols_df$grade_level=="Combination", na.rm=TRUE),
        sum(ols_df$charter=="Yes", na.rm=TRUE),
        sum(ols_df$charter=="No", na.rm=TRUE)),
  Percent = c(100*mean(ols_df$grade_level=="Elementary", na.rm=TRUE),
              100*mean(ols_df$grade_level=="Middle", na.rm=TRUE),
              100*mean(ols_df$grade_level=="High", na.rm=TRUE),
              100*mean(ols_df$grade_level=="Combination", na.rm=TRUE),
              100*mean(ols_df$charter=="Yes", na.rm=TRUE),
              100*mean(ols_df$charter=="No", na.rm=TRUE))
)

cat_t1 %>%
  gt(groupname_col = "Group", rowname_col = "Characteristic") %>%
  tab_header(title = "Campus sample: grade level and school type",
             subtitle = paste0("California K\u201312 schools, 2022 (N = ",
                               format(n_tot, big.mark=","), ")")) %>%
  fmt_number(columns = n, decimals = 0, sep_mark = ",") %>%
  fmt_number(columns = Percent, decimals = 1) %>%
  cols_label(n = "n", Percent = "%") %>%
  cols_align("right", columns = c(n, Percent)) %>%
  tab_style(style = cell_text(weight = "bold"), locations = cells_row_groups()) %>%
  tab_options(
    table.font.size = px(13),
    heading.title.font.size = px(15),
    heading.subtitle.font.size = px(12),
    heading.border.bottom.style = "none",
    column_labels.border.top.color = "grey60",
    column_labels.border.top.width = px(1),
    column_labels.border.bottom.color = "grey60",
    column_labels.border.bottom.width = px(1),
    column_labels.font.weight = "bold",
    row_group.border.top.style = "none",
    row_group.border.bottom.style = "none",
    table_body.hlines.color = "grey85",
    table_body.hlines.width = px(1),
    table.border.top.style = "none",
    table.border.bottom.color = "grey60",
    table.border.bottom.width = px(1),
    data_row.padding = px(9))

cont_t1 <- tibble::tibble(
  Measure = c("Chronic absenteeism (%)", "Campus NDVI",
              "Campus tree canopy (%)", "FRPM eligible (%)", "Enrollment (K\u201312)"),
  Mean   = c(mean(ols_df$chronic_rate,na.rm=TRUE), mean(ols_df$ndvi_campus,na.rm=TRUE),
             mean(ols_df$canopy_campus,na.rm=TRUE), mean(ols_df$pct_frpm_k12,na.rm=TRUE),
             mean(ols_df$enroll_k12,na.rm=TRUE)),
  SD     = c(sd(ols_df$chronic_rate,na.rm=TRUE), sd(ols_df$ndvi_campus,na.rm=TRUE),
             sd(ols_df$canopy_campus,na.rm=TRUE), sd(ols_df$pct_frpm_k12,na.rm=TRUE),
             sd(ols_df$enroll_k12,na.rm=TRUE)),
  Median = c(median(ols_df$chronic_rate,na.rm=TRUE), median(ols_df$ndvi_campus,na.rm=TRUE),
             median(ols_df$canopy_campus,na.rm=TRUE), median(ols_df$pct_frpm_k12,na.rm=TRUE),
             median(ols_df$enroll_k12,na.rm=TRUE)),
  IQR    = c(IQR(ols_df$chronic_rate,na.rm=TRUE), IQR(ols_df$ndvi_campus,na.rm=TRUE),
             IQR(ols_df$canopy_campus,na.rm=TRUE), IQR(ols_df$pct_frpm_k12,na.rm=TRUE),
             IQR(ols_df$enroll_k12,na.rm=TRUE)),
  Min    = c(min(ols_df$chronic_rate,na.rm=TRUE), min(ols_df$ndvi_campus,na.rm=TRUE),
             min(ols_df$canopy_campus,na.rm=TRUE), min(ols_df$pct_frpm_k12,na.rm=TRUE),
             min(ols_df$enroll_k12,na.rm=TRUE)),
  Max    = c(max(ols_df$chronic_rate,na.rm=TRUE), max(ols_df$ndvi_campus,na.rm=TRUE),
             max(ols_df$canopy_campus,na.rm=TRUE), max(ols_df$pct_frpm_k12,na.rm=TRUE),
             max(ols_df$enroll_k12,na.rm=TRUE))
)

cont_t1 %>%
  gt(rowname_col = "Measure") %>%
  tab_header(title = "Campus sample: outcome, exposures, and covariates",
             subtitle = paste0("California K\u201312 schools, 2022 (N = ",
                               format(n_tot, big.mark=","), ")")) %>%
  fmt_number(columns = c(Mean,SD,Median,IQR,Min,Max), decimals = 1) %>%
  fmt_number(columns = c(Mean,SD,Median,IQR,Min,Max), rows = Measure == "Campus NDVI", decimals = 2) %>%
  fmt_number(columns = c(Mean,SD,Median,IQR,Min,Max), rows = Measure == "Enrollment (K\u201312)",
             decimals = 0, sep_mark = ",") %>%
  cols_align("right", columns = c(Mean,SD,Median,IQR,Min,Max)) %>%
  tab_options(
    table.font.size = px(13),
    heading.title.font.size = px(15),
    heading.subtitle.font.size = px(12),
    heading.border.bottom.style = "none",
    column_labels.border.top.color = "grey60",
    column_labels.border.top.width = px(1),
    column_labels.border.bottom.color = "grey60",
    column_labels.border.bottom.width = px(1),
    column_labels.font.weight = "bold",
    table_body.hlines.color = "grey85",
    table_body.hlines.width = px(1),
    table.border.top.style = "none",
    table.border.bottom.color = "grey60",
    table.border.bottom.width = px(1),
    data_row.padding = px(9))

#wide version of categorical
n_tot <- nrow(ols_df)
pc <- function(cond) sprintf("%s (%.1f)", format(sum(cond, na.rm=TRUE), big.mark=","),
                             100*mean(cond, na.rm=TRUE))

## grade level (4 rows) on the left, school type (2 rows) on the right, padded to equal length so they sit side by side
cat_wide <- tibble::tibble(
  `Grade level`      = c("Elementary", "Middle", "High", "Combination"),
  `n (%)`            = c(pc(ols_df$grade_level=="Elementary"), pc(ols_df$grade_level=="Middle"),
                         pc(ols_df$grade_level=="High"), pc(ols_df$grade_level=="Combination")),
  `School type`      = c("Charter", "Traditional public", "", ""),
  `n (%) `           = c(pc(ols_df$charter=="Yes"), pc(ols_df$charter=="No"), "", "")
)

cat_wide %>%
  gt() %>%
  tab_header(title = "Campus sample: grade level and school type",
             subtitle = paste0("California K\u201312 schools, 2022 (N = ",
                               format(n_tot, big.mark=","), ")")) %>%
  cols_align("left", columns = c(`Grade level`, `School type`)) %>%
  cols_align("right", columns = c(`n (%)`, `n (%) `)) %>%
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_body(columns = c(`Grade level`, `School type`))) %>%
  tab_options(
    table.font.size = px(13),
    heading.title.font.size = px(15),
    heading.subtitle.font.size = px(12),
    heading.border.bottom.style = "none",
    column_labels.border.top.color = "grey60",
    column_labels.border.top.width = px(1),
    column_labels.border.bottom.color = "grey60",
    column_labels.border.bottom.width = px(1),
    column_labels.font.weight = "bold",
    table_body.hlines.color = "grey85",
    table_body.hlines.width = px(1),
    table.border.top.style = "none",
    table.border.bottom.color = "grey60",
    table.border.bottom.width = px(1),
    data_row.padding = px(9))

```

```{r}
## how many schools at the exact bounds, and what are they like?
ols_df %>%
  summarise(
    at_0    = sum(chronic_rate == 0,   na.rm = TRUE),
    at_100  = sum(chronic_rate == 100, na.rm = TRUE),
    near_0  = sum(chronic_rate < 1,    na.rm = TRUE),
    near_100= sum(chronic_rate > 99,   na.rm = TRUE))

## characteristics of the boundary schools -- are they small/atypical?
ols_df %>%
  filter(chronic_rate == 0 | chronic_rate == 100) %>%
  select(chronic_rate, enroll_k12, chronic_enroll, grade_level, charter) %>%
  arrange(desc(chronic_rate)) %>%
  as_tibble()%>%
  print(n = 40)

## enrollment distribution at the min/max vs overall
ols_df %>%
  mutate(bound = case_when(chronic_rate == 0 ~ "0%",
                           chronic_rate == 100 ~ "100%",
                           TRUE ~ "interior")) %>%
  group_by(bound) %>%
  summarise(n = n(),
            median_enroll = median(enroll_k12, na.rm = TRUE),
            median_elig   = median(chronic_enroll, na.rm = TRUE))
```

### Exposure correlation Check
```{r ndvi-canopy-correlation}
pairs <- tibble::tribble(
  ~scale,         ~ndvi_col,      ~canopy_col,
  "Campus",       "ndvi_campus",  "canopy_campus",
  "500m buffer",  "ndvi_500",     "canopy_500",
  "1000m buffer", "ndvi_1000",    "canopy_1000"
)

cor_check <- pairs |>
  rowwise() |>
  mutate(
    n        = sum(complete.cases(analysis[[ndvi_col]], analysis[[canopy_col]])),
    pearson  = cor(analysis[[ndvi_col]], analysis[[canopy_col]],
                   use = "complete.obs", method = "pearson"),
    spearman = cor(analysis[[ndvi_col]], analysis[[canopy_col]],
                   use = "complete.obs", method = "spearman")
  ) |>
  ungroup() |>
  select(scale, n, pearson, spearman)

cor_check

### Correlation matrix among vegetation exposures by spatial scale
cor_vars <- c(
  "ndvi_campus", "canopy_campus",
  "ndvi_500", "canopy_500",
  "ndvi_1000", "canopy_1000"
)

cor_lab <- c(
  ndvi_campus  = "NDVI (Campus)",
  canopy_campus = "Canopy % (Campus)",
  ndvi_500     = "NDVI (500m)",
  canopy_500   = "Canopy % (500m)",
  ndvi_1000    = "NDVI (1000m)",
  canopy_1000  = "Canopy % (1000m)"
)

# Spearman correlation matrix
veg_corr_mat <- analysis %>%
  select(all_of(cor_vars)) %>%
  cor(use = "pairwise.complete.obs", method = "spearman")

# Plot correlation matrix
corrplot::corrplot(
  veg_corr_mat,
  method = "color",
  type = "upper",
  order = "hclust",
  col = colorRampPalette(c("white", "#9ecae1", "#08519c"))(200),
  addCoef.col = "black",
  number.cex = 0.7,
  tl.col = "black",
  tl.cex = 0.8,
  diag = FALSE
)


```

### Table
```{r ols-table}
ols_results %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         scale   = factor(str_extract(Exposure, "Campus|\\d+m"),
                          levels = c("Campus","500m","1000m"))) %>%
  arrange(measure, scale) %>%
  select(Exposure, beta, ci_low, ci_high, p, N) %>%
  gt() %>%
  tab_header(title = "Campus greenspace (buffer exposures) and chronic absenteeism (OLS)",
             subtitle = "Coefficient per IQR; adjusted for FRPM, log-enrollment, charter; tract-clustered SEs") %>%
  fmt_number(c(beta, ci_low, ci_high), decimals = 3) %>% fmt_number(p, decimals = 3) %>%
  fmt_number(N, decimals = 0) %>%
  cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
  cols_label(ci_low = "95% CI", beta = "Coef (pp/IQR)") %>%
  tab_source_note("Negative = lower absenteeism with more greenness. Compare polygons vs 500m vs 1000m within each measure.") %>%
  tab_options(table.font.size = 11, column_labels.font.weight = "bold")
```


### Forest Plots
```{r forest-buffers, fig.width=9, fig.height=5}
ols_results %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         scale   = factor(str_extract(Exposure, "Campus|\\d+m"),
                          levels = c("Campus","500m","1000m")),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)")) %>%
  ggplot(aes(x = beta, y = scale, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  facet_wrap(~ measure, nrow = 1) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Campus greenspace and chronic absenteeism by spatial scale",
       subtitle = "OLS coefficient per IQR (95% CI); tract-clustered SEs; polygons vs 500m vs 1000m",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

```{r forest-buffers-ndvi-canopy, fig.width=7, fig.height=4}
## Grouped/combined figure for the REPORT: NDVI + tree canopy only (EVI removed here to fit the two-exposure space).
ols_results %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         scale   = factor(str_extract(Exposure, "Campus|\\d+m"),
                          levels = c("Campus","500m","1000m")),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)"),
         coef_lab = if_else(direction == "No association (CI spans 0)",
                            NA_character_, sprintf("(%.2f)", beta))) %>%
  filter(measure %in% c("NDVI", "Canopy")) %>%
  ggplot(aes(x = beta, y = scale, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.6) +
  geom_point(size = 2.5) +
  geom_text(aes(label = coef_lab), na.rm = TRUE, show.legend = FALSE,
            nudge_y = 0.28, size = 2.8) +
  facet_wrap(~ measure, nrow = 1) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Campus greenspace and chronic absenteeism by spatial scale",
       subtitle = "NDVI and tree canopy; OLS coefficient per IQR (95% CI); tract-clustered SEs",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

```{r exposure-histograms, fig.width=9, fig.height=6}
## Distributions of the (unscaled) exposures across schools, all three measures x three scales. Shows the spread each IQR-scaling is based on.
exp_cols <- c(ndvi_campus="NDVI Campus", ndvi_500="NDVI 500m", ndvi_1000="NDVI 1000m",
              evi_campus="EVI Campus",  evi_500="EVI 500m",  evi_1000="EVI 1000m",
              canopy_campus="Canopy Campus", canopy_500="Canopy 500m", canopy_1000="Canopy 1000m")

analysis %>%
  select(any_of(names(exp_cols))) %>%
  pivot_longer(everything(), names_to = "exp", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(exp = factor(exp_cols[exp], levels = unname(exp_cols))) %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 30, fill = "darkslateblue", color = "white", linewidth = 0.15) +
  facet_wrap(~ exp, scales = "free", ncol = 3) +
  labs(title = "Distribution of campus greenspace exposures",
       x = NULL, y = "Schools") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

```{r outcome-histogram, fig.width=6, fig.height=3.5}
## Distribution of the outcome (chronic absenteeism %) across schools.
ols_df %>%
  ggplot(aes(chronic_rate)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white", linewidth = 0.15) +
  labs(title = "Distribution of chronic absenteeism",
       x = "Chronic absenteeism %", y = "Schools") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"))
```

```{r model-output-histograms, fig.width=9, fig.height=3.5}
# Model-output distributions for a representative model (campus NDVI): fitted values and residuals, to inspect fit and the bounded-outcome behavior.
m_hist <- fit_ols_model <- lm(reformulate(c("ndvi_campus_iqr", covars), "chronic_rate"),
                              data = ols_df %>%
                                filter(if_all(all_of(c("chronic_rate","ndvi_campus_iqr",covars)), ~ !is.na(.))))

data.frame(fitted = fitted(m_hist), resid = resid(m_hist),
           std_resid = rstandard(m_hist)) %>%
  pivot_longer(everything(), names_to = "q", values_to = "value") %>%
  mutate(q = recode(q, fitted = "Fitted absenteeism %",
                    resid = "Residuals", std_resid = "Standardized residuals")) %>%
  ggplot(aes(value)) +
  geom_histogram(bins = 40, fill = "steelblue", color = "white", linewidth = 0.15) +
  facet_wrap(~ q, scales = "free", nrow = 1) +
  labs(title = "Campus NDVI model output", x = NULL, y = "Schools") +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"))
```

# 6. SES-quartile Stratified Analysis

```{r ses-stratified}
## Refit each exposure within each SES quartile. SES defines the stratum, so it is not adjusted within stratum (models keep FRPM, log-enrollment, charter).
fit_ols_stratum <- function(exposure, label, glevel) {
  need <- c("chronic_rate", exposure, covars, "GEOID")
  d <- ols_df %>% filter(ses_q == glevel, if_all(all_of(need), ~ !is.na(.)))
  if (nrow(d) < 30) return(NULL)
  m <- lm(reformulate(c(exposure, covars), "chronic_rate"), data = d)
  V <- vcovCL(m, cluster = d$GEOID)
  b <- coef(m)[exposure]; se <- sqrt(diag(V))[exposure]
  data.frame(Exposure = label, ses_q = glevel, beta = b,
             ci_low = b - 1.96*se, ci_high = b + 1.96*se,
             p = 2*pnorm(-abs(b/se)), N = nrow(d), row.names = NULL)
}

ses_levels <- levels(ols_df$ses_q)
ses_strat <- do.call(rbind, lapply(names(exp_map), function(e)
  do.call(rbind, lapply(ses_levels, function(g) fit_ols_stratum(e, exp_map[[e]], g)))))
```
### Table
```{r ses-strat-table}
ses_strat %>%
  mutate(ses_q = factor(ses_q, levels = ses_levels)) %>%
  arrange(Exposure, ses_q) %>%
  select(Exposure, ses_q, beta, ci_low, ci_high, p, N) %>%
  gt(groupname_col = "Exposure") %>%
  tab_header(title = "Campus greenspace and absenteeism, by SES quartile (OLS)",
             subtitle = "Coefficient per IQR within SES quartile; adjusted for FRPM, log-enrollment, charter; tract-clustered SEs") %>%
  fmt_number(c(beta, ci_low, ci_high), decimals = 3) %>% fmt_number(p, decimals = 3) %>%
  fmt_number(N, decimals = 0) %>%
  cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
  cols_label(ci_low = "95% CI", beta = "Coef (pp/IQR)", ses_q = "SES quartile") %>%
  cols_align("left", ses_q) %>%
  tab_source_note("Compare each exposure's coefficient across SES quartiles (effect modification by SES).") %>%
  tab_options(table.font.size = 9, column_labels.font.weight = "bold")
```

### Forest Plots
```{r forest-ses-strat1, fig.width=10, fig.height=7}
ses_strat %>%
  mutate(ses_q = factor(ses_q, levels = ses_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)")) %>%
  ggplot(aes(x = beta, y = ses_q, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by SES quartile",
       subtitle = "OLS coefficient per IQR (95% CI); tract-clustered SEs",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

```{r forest-ses-ndvi-canopy, fig.width=9, fig.height=6}
## REPORT version: NDVI + tree canopy only (six panels: 3 scales x 2 measures).
ses_strat %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         ses_q = factor(ses_q, levels = ses_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)"),
         coef_lab = if_else(direction == "No association (CI spans 0)",
                            NA_character_, sprintf("(%.2f)", beta))) %>%
  filter(measure %in% c("NDVI","Canopy")) %>%
  mutate(Exposure = factor(Exposure, levels = c(
    "NDVI Campus","NDVI 500m","NDVI 1000m","Canopy Campus","Canopy 500m","Canopy 1000m"))) %>%
  ggplot(aes(x = beta, y = ses_q, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  geom_text(aes(label = coef_lab), na.rm = TRUE, show.legend = FALSE,
            nudge_y = 0.28, size = 2.6) +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by SES quartile",
       subtitle = "NDVI and tree canopy; OLS coefficient per IQR (95% CI); tract-clustered SEs",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```
###Clustering Check

```{r se-clustering-check}
## Are observations independent? Schools sharing a census tract are not, because they share tract context (and, for tract-linked measures, exposure). quantifies the non-independence that justifies clustering SEs by tract GEOID.
cluster_overall <- ols_df %>%
  filter(!is.na(GEOID)) %>%
  group_by(GEOID) %>%
  summarise(schools_in_tract = n_distinct(CDSCode), .groups = "drop") %>%
  summarise(`Tracts (>=1 school)`      = n(),
            `Tracts w/ >1 school`      = sum(schools_in_tract > 1),
            `Schools in shared tracts` = sum(schools_in_tract[schools_in_tract > 1]),
            `Total schools`            = sum(schools_in_tract)) %>%
  mutate(`% schools sharing a tract` = 100 * `Schools in shared tracts` / `Total schools`)

cluster_overall %>%
  gt() %>%
  tab_header(title = "SE-clustering justification",
             subtitle = "Schools sharing a census tract are non-independent, justifying tract-clustered standard errors") %>%
  fmt_number(`% schools sharing a tract`, decimals = 1) %>%
  fmt_number(c(`Tracts (>=1 school)`, `Tracts w/ >1 school`,
               `Schools in shared tracts`, `Total schools`), decimals = 0) %>%
  tab_source_note("A high share of schools sharing tracts indicates non-independence; standard errors are clustered by tract GEOID throughout.") %>%
  tab_options(table.font.size = 11, column_labels.font.weight = "bold")
```

# 8. Grade Level Stratified Analysis

```{r grade-counts}
ols_df %>% filter(!is.na(grade_level)) %>%
  count(`Grade level` = grade_level) %>%
  gt() %>% tab_header(title = "Schools per grade level") %>%
  tab_options(table.font.size = 12, column_labels.font.weight = "bold")
```

```{r grade-stratified}
## Refit each exposure (all scales) within each grade band. Same covariates as the main models; SEs clustered by tract GEOID.
grade_levels <- levels(ols_df$grade_level)

fit_ols_grade <- function(exposure, label, glevel) {
  need <- c("chronic_rate", exposure, covars, "GEOID")
  d <- ols_df %>% filter(grade_level == glevel, if_all(all_of(need), ~ !is.na(.)))
  if (nrow(d) < 30) return(NULL)
  m <- lm(reformulate(c(exposure, covars), "chronic_rate"), data = d)
  V <- vcovCL(m, cluster = d$GEOID)
  b <- coef(m)[exposure]; se <- sqrt(diag(V))[exposure]
  data.frame(Exposure = label, grade = glevel, beta = b,
             ci_low = b - 1.96*se, ci_high = b + 1.96*se,
             p = 2*pnorm(-abs(b/se)), N = nrow(d), row.names = NULL)
}

grade_strat <- do.call(rbind, lapply(names(exp_map), function(e)
  do.call(rbind, lapply(grade_levels, function(g) fit_ols_grade(e, exp_map[[e]], g)))))

grade_strat %>%
  mutate(grade = factor(grade, levels = grade_levels)) %>%
  arrange(Exposure, grade) %>%
  select(Exposure, grade, beta, ci_low, ci_high, p, N) %>%
  gt(groupname_col = "Exposure") %>%
  tab_header(title = "Campus greenspace and absenteeism, by grade level (OLS)",
             subtitle = "Coefficient per IQR within grade band; adjusted for FRPM, log-enrollment, charter; tract-clustered SEs") %>%
  fmt_number(c(beta, ci_low, ci_high), decimals = 3) %>% fmt_number(p, decimals = 3) %>%
  fmt_number(N, decimals = 0) %>%
  cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
  cols_label(ci_low = "95% CI", beta = "Coef (pp/IQR)", grade = "Grade level") %>%
  cols_align("left", grade) %>%
  tab_source_note("Compare each exposure's coefficient across grade levels.") %>%
  tab_options(table.font.size = 9, column_labels.font.weight = "bold")
```

###Forest Plots

```{r forest-grade, fig.width=10, fig.height=8}
grade_strat %>%
  mutate(grade = factor(grade, levels = grade_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)")) %>%
  ggplot(aes(x = beta, y = grade, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by grade level",
       subtitle = "OLS coefficient per IQR (95% CI); tract-clustered SEs",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```

```{r forest-grade-ndvi-canopy, fig.width=9, fig.height=6}
## REPORT version: NDVI + tree canopy only.
grade_strat %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         grade = factor(grade, levels = grade_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)")) %>%
  filter(measure %in% c("NDVI","Canopy")) %>%
  mutate(Exposure = factor(Exposure, levels = c(
    "NDVI Campus","NDVI 500m","NDVI 1000m","Canopy Campus","Canopy 500m","Canopy 1000m"))) %>%
  ggplot(aes(x = beta, y = grade, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by grade level",
       subtitle = "NDVI and tree canopy; OLS coefficient per IQR (95% CI); tract-clustered SEs",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```


```{r forest-primary-report, fig.width=7.5, fig.height=4}
## PRIMARY report figure: main (unstratified) NDVI + canopy associations,across the three spatial scales.
ols_results %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         scale   = factor(str_extract(Exposure, "Campus|\\d+m"),
                          levels = c("Campus","500m","1000m"),
                          labels = c("Campus","500 m","1000 m")),
         measure = factor(measure, levels = c("NDVI","Canopy"),
                          labels = c("NDVI", "Tree canopy")),
         sig = ci_low > 0 | ci_high < 0) %>%
  filter(measure %in% c("NDVI","Tree canopy")) %>%
  ggplot(aes(x = beta, y = scale)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey55", linewidth = 0.4) +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.18,
                 linewidth = 0.7, color = "grey30") +
  geom_point(aes(fill = sig), shape = 21, size = 3.2, color = "grey20", stroke = 0.5) +
  facet_wrap(~ measure) +
  scale_fill_manual(values = c(`TRUE` = "#1B7837", `FALSE` = "white"),
                    labels = c(`TRUE` = "Significant (95% CI excludes 0)",
                               `FALSE` = "Not significant"), name = NULL) +
  labs(title = "Campus greenspace and chronic absenteeism",
       subtitle = "Adjusted OLS coefficient per IQR increase; tract-clustered 95% CI",
       x = "Change in chronic absenteeism (percentage points) per IQR",
       y = NULL) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold"),
        strip.text = element_text(face = "bold"),
        panel.grid.minor = element_blank(),
        legend.position = "bottom")
```
### Clustering Check
```{r grade-clustering-check}
## Re-check the clustering justification WITHIN each grade stratum. do schools still share tracts once split by grade? 
ols_df %>%
  filter(!is.na(grade_level)) %>%
  group_by(grade_level, GEOID) %>%
  summarise(schools_in_tract = n_distinct(CDSCode), .groups = "drop") %>%
  group_by(`Grade level` = grade_level) %>%
  summarise(`Tracts (>=1 school)`      = n(),
            `Tracts w/ >1 school`      = sum(schools_in_tract > 1),
            `Schools in shared tracts` = sum(schools_in_tract[schools_in_tract > 1]),
            `Total schools`            = sum(schools_in_tract),
            .groups = "drop") %>%
  mutate(`% schools sharing a tract` = 100 * `Schools in shared tracts` / `Total schools`) %>%
  gt() %>%
  tab_header(title = "SE-clustering check within grade strata",
             subtitle = "Do schools still share tracts after splitting by grade level?") %>%
  fmt_number(`% schools sharing a tract`, decimals = 1) %>%
  fmt_number(c(`Tracts (>=1 school)`, `Tracts w/ >1 school`,
               `Schools in shared tracts`, `Total schools`), decimals = 0) %>%
  cols_align("left", `Grade level`) %>%
  tab_source_note("Higher % sharing = tract clustering remains justified within that grade stratum. Note within-grade sharing is typically lower than overall, since same-tract schools are often different levels (an elementary + a high school).") %>%
  tab_options(table.font.size = 10, column_labels.font.weight = "bold")
```
# 9. Interaction 

```{r ols-interaction}
## Effect modification: does the greenspace-absenteeism association differ by SES? Add a multiplicative exposure x SES_INDEX term to the full-sample OLS. The interaction coefficient tests effect modification (SES continuous). Tract-clustered SEs, same covariates as the main models.

fit_ols_interaction <- function(exposure, label) {
  int_term <- paste0(exposure, ":SES_INDEX")
  need <- c("chronic_rate", exposure, "SES_INDEX", covars, "GEOID")
  d <- ols_df %>% filter(if_all(all_of(need), ~ !is.na(.)))

  ## full model: exposure * SES = exposure + SES + exposure:SES  (+ covariates)
  f <- reformulate(c(paste0(exposure, " * SES_INDEX"), covars), "chronic_rate")
  m <- lm(f, data = d)
  V <- vcovCL(m, cluster = d$GEOID)
  ct <- lmtest::coeftest(m, vcov. = V)

  b  <- ct[int_term, "Estimate"]
  se <- ct[int_term, "Std. Error"]
  data.frame(Exposure = label,
             interaction = b,
             ci_low = b - 1.96*se, ci_high = b + 1.96*se,
             p = 2*pnorm(-abs(b/se)), N = nrow(d), row.names = NULL)
}

ols_interaction <- do.call(rbind, lapply(names(exp_map), function(e)
  fit_ols_interaction(e, exp_map[[e]])))
```
### Table
```{r ols-interaction-table}
ols_interaction %>%
  mutate(measure = str_remove(Exposure, " (Campus|\\d+m)$"),
         scale   = factor(str_extract(Exposure, "Campus|\\d+m"),
                          levels = c("Campus","500m","1000m"))) %>%
  arrange(measure, scale) %>%
  select(Exposure, interaction, ci_low, ci_high, p, N) %>%
  gt() %>%
  tab_header(title = "Greenspace \u00d7 SES interaction (effect modification test)",
             subtitle = "OLS interaction coefficient (exposure \u00d7 SES_INDEX); tract-clustered SEs; adjusted for FRPM, log-enrollment, charter") %>%
  fmt_number(c(interaction, ci_low, ci_high), decimals = 4) %>%
  fmt_number(p, decimals = 3) %>% fmt_number(N, decimals = 0) %>%
  cols_merge(c(ci_low, ci_high), pattern = "[{1}, {2}]") %>%
  cols_label(ci_low = "95% CI", interaction = "Interaction coef", p = "p") %>%
  tab_source_note("Interaction coef = change in the per-IQR greenspace slope per one-unit increase in SES_INDEX. p < 0.05 indicates significant effect modification by SES.") %>%
  tab_options(table.font.size = 10, column_labels.font.weight = "bold")
```

```{r forest-ses-strat2, fig.width=10, fig.height=7}
#full strat interaction plot

## p-int per Exposure (facet) from the OLS interaction models
p_df <- ols_interaction %>%
  mutate(p_lab = paste0("p-int = ", formatC(p, format = "f", digits = 3))) %>%
  select(Exposure, p_lab)

ses_strat %>%
  mutate(ses_q = factor(ses_q, levels = ses_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)"),
         ## coefficient label: blank for null; else (x.xx), placed past the CI end
         coef_lab  = ifelse(direction == "No association (CI spans 0)", "",
                            sprintf("(%.2f)", beta)),
         lab_x     = ifelse(beta > 0, ci_high, ci_low),
         lab_hjust = ifelse(beta > 0, -0.15, 1.15)) %>%
  ggplot(aes(x = beta, y = ses_q, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  geom_text(aes(x = lab_x, label = coef_lab, hjust = lab_hjust),
            size = 2.5, show.legend = FALSE) +
  geom_text(data = p_df, aes(x = Inf, y = Inf, label = p_lab),
            inherit.aes = FALSE, hjust = 1.05, vjust = 1.4, size = 2.8, color = "grey25") +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by SES quartile",
       subtitle = "OLS coefficient per IQR (95% CI); tract-clustered SEs; p-int = greenspace \u00d7 SES interaction p",
       x = "Coefficient: pp change in absenteeism per IQR", y = NULL) +
  scale_x_continuous(expand = expansion(mult = 0.22)) +
  scale_y_discrete(expand = expansion(add = c(0.5, 1.2))) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```


```{r forest-ses-strat, fig.width=10, fig.height=7}
#removing EVI for Report
# p-int per Exposure from the OLS interaction models — NDVI & canopy only
p_df <- ols_interaction %>%
  filter(!str_detect(Exposure, "EVI")) %>%
  mutate(p_lab = paste0("p-int = ", formatC(p, format = "f", digits = 3))) %>%
  select(Exposure, p_lab)

ses_strat %>%
  filter(!str_detect(Exposure, "EVI")) %>%
  mutate(ses_q = factor(ses_q, levels = ses_levels),
         direction = case_when(ci_high < 0 ~ "Protective (coef < 0)",
                               ci_low  > 0 ~ "Adverse (coef > 0)",
                               TRUE ~ "No association (CI spans 0)"),
         coef_lab  = ifelse(direction == "No association (CI spans 0)", "",
                            sprintf("(%.2f)", beta)),
         lab_x     = ifelse(beta > 0, ci_high, ci_low),
         lab_hjust = ifelse(beta > 0, -0.15, 1.15)) %>%
  ggplot(aes(x = beta, y = ses_q, color = direction)) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "grey40") +
  geom_errorbarh(aes(xmin = ci_low, xmax = ci_high), height = 0.25, linewidth = 0.55) +
  geom_point(size = 2) +
  geom_text(aes(x = lab_x, label = coef_lab, hjust = lab_hjust),
            size = 2.5, show.legend = FALSE) +
  geom_text(data = p_df, aes(x = Inf, y = Inf, label = p_lab),
            inherit.aes = FALSE, hjust = 1.05, vjust = 1.4, size = 2.8, color = "grey25") +
  facet_wrap(~ Exposure, ncol = 3) +
  scale_color_manual(values = c("Protective (coef < 0)" = "darkolivegreen",
                                "Adverse (coef > 0)"     = "#B2182B",
                                "No association (CI spans 0)" = "grey55"), name = NULL) +
  labs(title = "Greenspace and absenteeism by SES quartile",
       subtitle = "OLS coefficient per IQR (95% CI); tract-clustered SEs; p-int = greenspace \u00d7 SES interaction p",
       x = "Coefficient: IQR change is associated with a pp change in absenteeism", y = NULL) +
  scale_x_continuous(expand = expansion(mult = 0.22)) +
  scale_y_discrete(expand = expansion(add = c(0.5, 1.2))) +
  theme_minimal(base_size = 10) +
  theme(plot.title = element_text(face = "bold"), legend.position = "top")
```






